For submission to the Journal of Integrated Pest Management

### 2021-05-21 ofv_fife_2021.Rmd
Word smithing up the distinctions between virus detection methods to clarify exactly what happened, where and to what plants.

We report outbreaks of *Orchid fleck dichorhavirus* (OFV-Orc) infecting three new hosts: *Liriope* spp. *Ophiopogon* and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in Leon and Alachua Counties, Florida. Strains of OFV-Orc infect over 50 plant species belonging to the plant families Orchidaceae, Asparagaceae (Nolinoidaea), and infects *Citrus* (Rutaceae), as citrus leprosis disease. The only known vectors of OFV-Orc are flat mites from the genus *Brevipalpus* (Trombidiformes: Tenuipalpidae). Florida has various native and introduced plants in the landscape which *Brevipalpus* spp. feed on, which are potentially susceptible to OFV-Orc.

Chlorotic ringspots flecking were seen affecting Liriopogons (*Liriope* spp. and *Ophiopogon* spp.) in Leon County, FL. Nearby *A. elatior* also appeared chlorotic. Local diagnostics returned negative for common plant viruses, so samples were sent to the Florida Department of Agriculture and Consumer Services for identification. Orchid-infecting *Orchid fleck dichorhavirus* was confirmed via conventional reverse transcription polymerase chain reaction of L-gene (RNA2) and Sanger sequencing of amplicons. Amplicons had a 98% identity with known OFV sequences available in NCBI GenBank.

New samples from symptomatic plants were collected in both Leon and Alachua counties. High-throughput sequencing (HTS) at the USDA-ARS detected OFV-Orc and OFV-Orc2 in *Aspidistra* and *Liriope*, respectively. Three potential mite vectors were identified via cryo-scanning electron microscopy (Cryo-SEM): *Brevipalpus californicus* (Banks) sensu lato, *B. obovatus* Donnadieu, and *B. confusus* Baker.

@@

### Keywords:
False spider mite, flat mite, *Brevipalpus*-transmitted viruses, *Liriope*, Nolinoideae, *Ophiopogon*, Ruscaceae, Rutaceae, Asparagaceae, orchid, Orchidaceae, pests, ornamental plants, orchid fleck virus.

@@

OFV-infected plants exhibit various symptoms depending on the infected plant species as well as the strain of the OFV associated with the infection [@Kubo2009b], but symptoms typically appear as chlorotic flecks, which ultimately coalesce into larger spots or ringspot patterns<!--(`r citef("fig_1")`, `r citef("fig_2")`)-->.

@@

The initial site of collection was visited two more times during 2020 to gather plants for identification of the unidentified plant pathogen. These surveys were conducted during July and August to collect more putatively infected plants, including more *Liriope* spp. as well as a new species which also belongs to the family Asparagaceae; *Aspidistra elatior* Blume, which was suspected to be infected due to unusually chlorotic leaves (`r citef("fig_2")`). Upon collection, these new samples were sent to the Florida Department of Agriculture and Consumer Services (FDACS) for identification.

The presence of OFV was confirmed using OFV generic R2-Dicho-GF and R2-Dicho-GR primers by one step conventional RT-PCR [@Roy2020], amplifying ~800 nt of L-gene (RNA2) amplicon from an infected *Liriope* leaf sample. Sanger sequencing of RT-PCR amplicons shared 98\% nucleotide identity with orchid strains of OFV: OFV-Orc1 and OFV-Orc2 (GenBank Accession numbers: AB244418 and LC222630) [@Kondo2006; @Kondo2017].

@@

OFV has been reported in other Nolinoidaea in Australia [@Dietzgen2018b; @Mei2016], including *Liriope spicata* (Thunb.) Lour. [@Mei2016] but not in the US. The Florida collected plants of *Liriope* spp., cv 'Gigantea' are thought to belong to either *Liriope muscari* or *Liriope gigantea*. We are not aware of any previous report of OFV infection in *Ophiopogon* plants. Although @Zheng2013 mention the association of *B. californicus* with *A. elatior*, but never reported OFV symptoms in this plant. However, our finding will be notified as the first report of OFV in Florida, in the US on ornamentals and among them, *A. elatior* is a new natural host of OFV.

### Conclusion
The dichorhavirus that infects citrus in Hawaii, Mexico, Colombia, and South Africa are identical to the OFV in gene order, content, and the genome sequence. According to the International Committee on Taxonomy of Viruses (ICTV) classification, OFV consist of two orchid strains (OFV-Orc1 and OFV-Orc2) and two citrus strains (OFV-Cit1 and OFV-Cit2). Both the orchid strains of OFV infects citrus [@Roy2020], but none of the citrus strains have been reported from any orchid species. Detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivating *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and eradicated in the mid-1960s. In fact, @Kitajima2011a found that the Citrus Leprosis virus (CiLV) which previously affected Florida citrus was a nuclear type of citrus leprosis closely related to OFV strains [@Roy2020; @Hartung2015]. Association of a distant relative of OFV named Citrus leprosis dichorhavirus-N0 (CiLV-N0) was confirmed in relation to the leprosis disease outbreak in Florida [@Hartung2015]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the threat of different strains of OFV on citrus; which will be a definite concern to the US multi-billion dollar citrus industry. *B. californicus*, as well as *B. yothersi* (Baker), are both known vectors of Dichorhaviruses (OFV) [@Kondo2003; @Beltran-Beltran2020] and *B. obovatus* is a suspected vector as well [@Childers2003]. All three mite species/complexes are present in Florida [@Childers2003; @Akyazi2017] (`r citef("fig_4")`). Therefore, it is critical to identify the vector of OFVs in Florida and monitor its spread to determine the risk this virus represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.

@@

|Scientific Name|Common Names|Symptoms Observed|County|
|:-----------|:--------------|:----------------|:-----|
|*Liriope muscari* Bailey| Lilyturf, Orchardgrass, Monkeygrass| Ringspots, Chlorotic Flecking, Necrotic Lesions|Alachua & Leon|
|*Liriope gigantea*\* Hume| Giant Lilyturf| Ringspots, Chlorotic Flecking, Necrotic Lesions|Alachua & Leon|
|*Ophiopogon japonicus* Ker Gawl.| Dwarf Lilyturf, Mondo Grass, Snake's beard| Ringspots, Chlorotic Flecking, Necrotic Lesions|Leon|
|*Ophiopogon intermedius*\*\* Don| Aztec Grass, 'Argenteomarginatus' | Ringspots, Chlorotic Flecking, Necrotic Lesions|Alachua & Leon|
|*Aspidistra elatior* Blume| Cast Iron Plant, Bar-room Plant | Chlorosis, Chlorotic Flecking, Necrotic Lesions|Leon|
`r tab_nums(name = "tab_1", caption = "* *L. gigantea* has been traditionally classified as seperate from *L. muscari* by @Broussard2007 and @Fantz2015, although this distinction has been challenged by @Wang2014 and @Masiero2020. * * *O. intermedius* is sometimes misclassified as *Liriope muscari* ‘Variegated Evergreen Giant’ [@Fantz2008b, @Fantz2009] or ‘Grandiflora White’ [@Fantz2009].")`

>>>>

We describe the first outbreaks of *Orchid fleck dichorhavirus*, belonging to the orchid-infecting subgroup (OFV-Orc), from three unreported hosts: *Liriope muscari* (Decaisne) Bailey, *Ophiopogon intermedius* Don and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in Leon and Alachua Counties, FL. Strains of OFV-Orc infect over 50 plant species belonging to the plant families Orchidaceae, Asparagaceae (Nolinoidaea), and infects *Citrus* (Rutaceae), as citrus leprosis disease. The only known vectors of OFV-Orc are flat mites from the genus *Brevipalpus* (Trombidiformes: Tenuipalpidae). Florida has various native and introduced plants in the landscape which *Brevipalpus* spp. feed on, which are potentially susceptible to OFV-Orc.

Chlorotic ringspots and flecking were seen affecting Liriopogons (*Liriope* spp. and *Ophiopogon* spp.) in Leon County, FL. Adjacent *A. elatior* also appeared chlorotic. Local diagnostics returned negative for common plant pathogens, therefore samples were sent to the Florida Department of Agriculture and Consumer Services for identification. OFV-Orc was confirmed in *L. muscari* and *A. elatior* via conventional reverse transcription polymerase chain reaction and Sanger sequencing of amplicons. Amplicons had a 98% identity with known sequences of Orchid fleck virus available in NCBI GenBank.

New samples from symptomatic plants were collected in both Leon and Alachua counties and sent to the USDA-ARS. Conventional RT-PCR, RT-qPCR and Sanger sequencing confirmed the presence of Orchid fleck virus, OFV-Orc subgroup. High-throughput sequencing detected OFV-Orc and OFV-Orc2 in *A. elatior* and *L. muscari*, respectively. Three potential mite vectors were identified via cryo-scanning electron microscopy (Cryo-SEM): *Brevipalpus californicus* (Banks) sensu lato, *B. obovatus* Donnadieu, and *B. confusus* Baker.

@@

### Keywords:
False spider mite, flat mite, *Brevipalpus*-transmitted viruses, *Liriope*, Nolinoideae, *Ophiopogon*, *Aspidistra*, Ruscaceae, Rutaceae, Asparagaceae, orchid, Orchidaceae, pests, ornamental plants, Orchid fleck virus.

@@

OFV-infected plants exhibit various symptoms depending on the infected plant species as well as the strain of the OFV associated with the infection [@Kubo2009b], but symptoms typically appear as chlorotic flecks, which ultimately coalesce into larger spots or ringspot patterns <!--(`r citef("fig_1")`, `r citef("fig_2")`)-->.

@@

The initial site of collection was visited two more times during 2020 to gather plants for identification of the unidentified plant pathogen. These surveys were conducted during July and August to collect more putatively infected plants, including more *Liriope* spp. as well as a new species which also belongs to the family Asparagaceae; *Aspidistra elatior* Blume, which was suspected to be infected due to unusually chlorotic leaves (`r citef("fig_2")`). Upon collection, these new samples were sent to the Florida Department of Agriculture and Consumer Services (FDACS) for identification. 
The FDACS determined that the virus was *Orchid fleck dichorhavirus* using previously published primers and methods [@Kubo2006a; @Kubo2006b; @Kubo2009a; @Kubo2009b; @RamosGonzalez2015]. Orchid subgroup OFV-Orc was identified following the methods described in @Kondo2017.

Samples sent to the USDA confirmed the presence of OFV via one step conventional RT-PCR and RT-qPCR, using generic R2-Dicho-GF and R2-Dicho-GR primers, amplifying ~800 nt of L-gene (RNA2) amplicon [@Roy2020]. Follow-up Sanger sequencing demonstrated a shared 98\% nucleotide identity with the orchid strain subgroup, OFV-Orc (isolates So and Br with GenBank Accession numbers: AB244418 and MK522807, respectively) [@Kondo2006; @Kondo2017]. High Throughput sequencing results confirmed the presence of OFV-Orc2 strain in the *Liriope* sample.

@@

OFV has been reported in other Nolinoidaea in Australia [@Dietzgen2018b; @Mei2016], including *Liriope spicata* (Thunb.) Lour. [@Mei2016] but not in the US. The *Liriope* spp., cv 'Gigantea' is considered a synonym of *Liriope muscari* [@Masiero2020; @Wang2014], but some consider it a separate species: *Liriope gigantea* Hume [@Fantz2015]. We are not aware of any previous report of OFV infection in *Ophiopogon* plants. Although @Zheng2013 mention the association of *B. californicus* with *A. elatior*, but never reported OFV symptoms in this plant. However, our finding will be notified as the first report of OFV in Florida, in the US on ornamentals and among them, *A. elatior* is a new natural host of OFV.

### Conclusion
The dichorhavirus that infects citrus in Hawaii, Mexico, Colombia, and South Africa are identical to the OFV in gene order, content, and the genome sequence. According to the International Committee on Taxonomy of Viruses (ICTV) classification, OFV consist of two orchid strains (OFV-Orc1 and OFV-Orc2) and two citrus strains (OFV-Cit1 and OFV-Cit2). Both the orchid strains of OFV infects citrus [@Roy2020], but none of the citrus strains have been reported from any orchid species. Detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivating *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and eradicated in the mid-1960s. In fact, @Kitajima2011a found that the Citrus Leprosis virus (CiLV) which previously affected Florida citrus was a nuclear type of citrus leprosis closely related to OFV strains [@Roy2020; @Hartung2015]. Association of a distant relative of OFV named Citrus leprosis dichorhavirus-N0 (CiLV-N0) was confirmed in relation to the leprosis disease outbreak in Florida [@Hartung2015]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the threat of different strains of OFV on citrus; which will be a definite concern to the US multi-billion dollar citrus industry. *B. californicus*, as well as *B. yothersi* (Baker), are both known vectors of Dichorhaviruses (OFV) [@Kondo2003; @Beltran-Beltran2020] and *B. obovatus* is a suspected vector as well [@Childers2003]. All three mite species/complexes are present in Florida [@Childers2003; @Akyazi2017] (`r citef("fig_4")`). Therefore, it is critical to identify the vector of these strains of OFV in Florida to monitor their spread, and to determine the risk they represent for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.

@@

|Scientific Name|Common Names|Symptoms Observed|County|
|:-----------|:--------------|:----------------|:-----|
|*Liriope muscari* (Decaisne) Bailey| Lilyturf, Orchardgrass, Monkeygrass| Ringspots, Chlorotic Flecking, Necrotic Lesions|Alachua & Leon|
|*Liriope gigantea*\* Hume| Giant Lilyturf| Ringspots, Chlorotic Flecking, Necrotic Lesions|Alachua & Leon|
|*Ophiopogon japonicus* Ker Gawl.| Dwarf Lilyturf, Mondo Grass, Snake's beard| Ringspots, Chlorotic Flecking, Necrotic Lesions|Leon|
|*Ophiopogon intermedius*\*\* Don| Aztec Grass, 'Argenteomarginatus' | Ringspots, Chlorotic Flecking, Necrotic Lesions|Alachua & Leon|
|*Aspidistra elatior* Blume| Cast Iron Plant, Bar-room Plant | Chlorosis, Chlorotic Flecking, Necrotic Lesions|Leon|
`r tab_nums(name = "tab_1", caption = "* *L. gigantea* has been traditionally classified as seperate from *L. muscari* by @Broussard2007 and @Fantz2015, although this distinction has been challenged by @Wang2014 and @Masiero2020. * * *O. intermedius* is sometimes misclassified as *Liriope muscari* ‘Variegated Evergreen Giant’ [@Fantz2008b, @Fantz2009] or ‘Grandiflora White’ [@Fantz2009].")`
